---
title: "Critical Care Delayed Discharges"
author: "AJ Clarkin"
date: "`r format(Sys.Date(), '%d/%m/%Y')`"
output:
  pdf_document: default
  html_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```




```{r environment}

library(tidyverse)
library(here)
library(lubridate)
library(scales)
# library(patchwork)
library(viridisLite)


datafile = "data_processed/data.csv"

```


```{r import-data}

# Import the data
# This data as already been processed (column names corrected, dates converted)

df <- read_csv(here(datafile))


```



```{r report-settings}

# Find the latest date in the data file and use that as the latest month to report on.
# Variables are below to override.

max_date <- max(df$dc_isodate)
publish_year <- year(max_date)
publish_month <- month(max_date)


# To override
# publish_year = 2022
# publish_month = 3



# We know the latest month we want to report on, now we need to find the earliest (to ensure we only show 12 months
# which is required from the levels within the month factor).

start_month <- 0

if(publish_month == 12) {
    start_month <- 1
    start_year <- publish_year
} else {
    start_month <- publish_month + 1
    start_year <- publish_year - 1
}


# Now use these to create a text date range of the form "May 2021 - April 2022" for use in graphics
txt_daterange <- paste(month.name[start_month], start_year, "-", month.name[publish_month], publish_year)

# Build the start of the filepath for exporting images
# Image content type will be appended
txt_imgfilename <- paste(here("images"), "/", publish_year, "-", str_pad(publish_month, 2, pad="0"), "_", sep="")
    


```



## Source Data

Data was taken from WardWatcher.

Date / time of admission and ready for discharge were used to calculate treatment time.

Date / time of ready for discharge and actual discharge were used to calculate delay.

The national standard is that discharge must occur within 4 hours from the decision and the SICSAG national audit uses the 4 hours as a cutoff to identify patients who have had a delayed discharge. That cutoff is used here.




```{r filter-data}

# Remove any of the data from before the time period we are interested in

df <- df %>% 
    filter(!(dc_year < start_year)) %>% 
    filter(!(dc_year == start_year & dc_month < start_month))

# Drop some of the columns

df <- df %>% 
    select(dc_isodate, dc_year, dc_month, treat_mins, delay_mins, total_mins, dest, dest_name, dc_reason)

# And now filter out the self-discharges before doing the main processing.
# This will ensure the denominator used excludes these.
# df <- df %>%  
#     filter(dc_reason != "H. Self discharge")

```





```{r grouping}

## Find all the patients who went to ward or HDU, classify as delayed or not, and then count the whole number of bed days they were delayed.
# So is_delayed looks at 4h+ delays
#   is_delayed_day counts the whole days they were delayed.

df_delaydays <- df %>%
    mutate(month = factor(dc_month,
                          levels = c(start_month:12, 1:publish_month),
                          labels = c(month.abb[start_month:12], month.abb[1:publish_month])),
           is_delayed = if_else(delay_mins <= 240, FALSE, TRUE),
           is_delayed_day = if_else(floor(delay_mins / 1440) > 0, TRUE, FALSE)
           )

# Summary Table 
# df_delaydays %>% rename(Month = month) %>%
#     group_by(Month) %>% 
#     summarise(Patients = n())

```





```{r summary-table}

# Total patients by delay status or not per month

df_monthsummary <- df_delaydays %>% 
    filter(dest != "B. Intensive Care Unit" & dest != "Z. Mortuary") %>%
    group_by(month, is_delayed) %>% 
    summarise(grp_count=n(),
              grp_delay_mins = sum(delay_mins),
              grp_treat_mins = sum(treat_mins, na.rm = TRUE),
              grp_total_mins = sum(total_mins, na.rm= TRUE)
              ) %>% 
    
    group_by(month) %>% 
    summarise(is_delayed, grp_count,
              month_count=sum(grp_count),
              grp_pc = grp_count / month_count,
              grp_delay_mins, grp_treat_mins, grp_total_mins,
              tot = sum(grp_total_mins)
              ) 


```

## Delayed Discharges

The following charts look at discharges to destinations within NHS Grampian. Transfers to Critical Care Units in other health boards
were excluded as these make up a small number, are not modifiable, and are usually timed based on staffing, ambulance availability, and capacity at the receiving hospital.

Patients who died within Critical Care were also excluded as they would not be listed for discharge.


```{r unit-charts, fig.width=8}

# Connected scatterplot of rate of delayed discharge.

p1 <- df_monthsummary %>% filter(is_delayed == TRUE) %>%
    ggplot(aes(x=month, y=grp_pc, group=1)) +
    geom_line(colour="light blue") +
    geom_point(colour = "blue") +
    theme_bw() +
    ggtitle("Delayed Discharges from Critical Care - Percentage of Discharges", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Percentage of All Discharges") +
    scale_y_continuous(labels = percent, limits = c(0,1))



# Bar chart showing total number of patients delayed each month.

p2 <- df_monthsummary %>% filter(is_delayed == TRUE) %>%
    ggplot(aes(month, grp_count), ) +
    geom_bar(stat = "identity", fill="light blue") +
    # scale_y_continuous(breaks = seq(0, 25, 4)) +
    # scale_x_discrete(breaks = NULL) +
    theme_light() +
    ggtitle("Delayed Discharges from Critical Care - Number of Patients", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Number of Patients Delayed")

p1
p2
```




## Delayed Discharges By Destination

WardWatcher records the location the patient was actually discharged to. So, for some of the
patients who have been delayed the destination has changed and they have gone directly home
rather than to the ward they were listed for. Patients able to go home directly without requirement
for a step-down bed would not usually experience a delay.


```{r, fig.width=8}

df_destsummary <- df_delaydays %>% 
    filter(dest != "B. Intensive Care Unit" & dest != "Z. Mortuary") %>%
    group_by(month, dest, is_delayed) %>% 
    summarise(grp_count=n(),
              grp_delay_mins = sum(delay_mins),
              grp_treat_mins = sum(treat_mins, na.rm = TRUE),
              grp_total_mins = sum(total_mins, na.rm= TRUE)
              ) %>% 
    
    group_by(month) %>% 
    summarise(dest, is_delayed, grp_count,
              month_count=sum(grp_count),
              grp_pc = grp_count / month_count,
              grp_delay_mins, grp_treat_mins, grp_total_mins,
              month_tot_mins = sum(grp_total_mins)
              ) 

# df_destsummary



```

```{r fig.width=8}

# Stacked bar chart of no of delayed patients by month showing discharge destination

df_destsummary %>% filter(is_delayed == TRUE) %>% 
    ggplot(aes(x=month, y=grp_count, fill=dest)) +
    geom_bar(stat="identity", position = position_stack(reverse = TRUE)) +
    theme_light() +
    ggtitle("Discharge Destination for Patients with Delayed Discharge", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Number of Patients")




# Connected line for total number of delays by month

df_destsummary %>% filter(is_delayed == TRUE) %>% 
    group_by(month) %>% 
    summarise(delay_days = sum(grp_delay_mins) / (60*24)) %>% 
    ggplot(aes(x=month, y=delay_days, group=1)) +
    geom_point(colour="dark blue") +
    geom_line(colour="blue") +
    theme_light() +
    ggtitle("Total Duration of Delay in Critical Care", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Cumulative Delay (days)")



# Connected line for number of delays by month by destination

df_destsummary %>% filter(is_delayed == TRUE) %>% 
    # group_by(month) %>% 
    # summarise(delay_days = sum(grp_delay_mins) / (60*24)) %>% 
    ggplot(aes(x=month, y=(grp_delay_mins / (60*24)), colour=dest, group=dest)) +
    geom_point() +
    geom_line() +
    theme_light() +
    ggtitle("Total Duration of Delay in Critical Care by Destination", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Cumulative Delay (days)") +
    labs(colour = "Destination Type")
```
\newpage

## Ward Discharges

The greatest proportion of delays are to wards. Here the wards are grouped according to ward numbers.




```{r ward-type, fig.width=8}

df_ward <- df_delaydays %>% 
    filter(dest == "A. Ward" & grepl("Ward", dest_name)) %>% 
    mutate(ward_group = as.numeric(substring(dest_name, first=6, last=6)) *100,
           ward_group = as.factor(ward_group)
           )


df_ward %>% filter(is_delayed == TRUE) %>% 
    group_by(month, ward_group) %>% 
    summarise(group_delay_days = sum(delay_mins) / (60 * 24)) %>% 
    ggplot(aes(x=month, y=group_delay_days, colour=ward_group, group=ward_group)) +
    geom_point() +
    geom_line() +
    theme_light() +
    ggtitle("Total Duration of Delay in Critical Care by Ward Group", subtitle = txt_daterange) +
    xlab("Month") +
    ylab("Cumulative Delay (days)") +
    labs(colour = "Ward Grouping")

```




