---
title: "ImportDischargeData"
author: "AJ Clarkin"
date: "25/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Import Discharge Data From WardWatcher

Separate the initial import and processing of data from the report generation.

Data from ward watcher should have the following columns:
- Discharged on (date)
- Discharged at (time)
- Gap between ready & discharge (mins)
- Gap considered
- Destination (type)
- Reason discharged
- Calendar day stay


Data should be saved into the data_raw folder.


```{r imports}

library(tidyverse)
library(here)
library(lubridate)

filename_out = "data_processed/data.csv"

filename_in = "data_raw/may22.tsv"


```




```{r import-data}

df_data <- read_csv(here(filename_out))


df_new <- read_tsv(here(filename_in),
                   skip = 1,
                   col_names = c("dc_date", "dc_time", "delay_mins", "gap_considered", "dest", "dc_reason", "stay_caldays"))


df_new <- df_new %>% 
    mutate(dc_isodate = dmy(dc_date),
           dc_year = year(dc_isodate),
           dc_month = month(dc_isodate)
           ) 



```



```{r de-duplicate}

# We have the previous saved data and the new data set.
# Merge them and then exclude duplicates.
# Print some info about this

print(paste("Rows in current data set: ", nrow(df_data)))
print(paste("Rows in new data set: ", nrow(df_new)))

df_merge <- bind_rows(df_data, df_new)

print(paste("After merging: ", nrow(df_merge)))

df_merge <- distinct(df_merge)
print(paste("Total number of distinct rows in merged dataframe: ",  nrow(df_merge)))

```




```{r export-data}

write_csv(df_merge, filename_out)


```

