---
title: "Unified Import"
author: "AJ Clarkin"
date: "25/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Note

This unifies the two previous versions and trims down the number of fields needed for import and export. Could also exclude delay_mins but when I run it the result 


### Import Discharge Data From WardWatcher

Separate the initial import and processing of data from the report generation.

Data from ward watcher should have the following columns:
- Discharged on (date)
- Discharged at (time)
- Gap between ready & discharge (mins)
- Destination (type)
- Destination (name)
- Reason discharged
- Date ready for discharge
- Time ready for discharge
- Date admitted to this unit
- Time admitted to this unit


Data should be saved into the data_raw folder.


```{r imports}

library(tidyverse)
library(here)
library(lubridate)

filename_out = "data_processed/data.csv"

filename_in = "data_raw/april23.tsv"


```




```{r import-data}

df_data <- read_csv(here(filename_out))


df_new <- read_tsv(here(filename_in),
                   skip = 1,
                   col_names = c("dc_date", "dc_time", "delay_mins", "dest", "dest_name", "dc_reason", 
                                 "rd_date", "rd_time", "ad_date", "ad_time" ))


# convert ready date/time and admit date/time to isodates
# then subtract to get actual treatment time in mins


df_new <- df_new %>% 
    mutate(dc_isodate = dmy_hms(paste(dc_date, dc_time)),
           dc_year = year(dc_isodate),
           dc_month = month(dc_isodate),
           ready_isodate = dmy_hms(paste(rd_date, rd_time)),
           admit_isodate = dmy_hms(paste(ad_date, ad_time)),
           treat_mins = as.numeric(ready_isodate - admit_isodate, units = "mins"),
           total_mins = as.numeric(dc_isodate - admit_isodate, units = "mins")
           ) 



```





```{r de-duplicate}

# We have the previous saved data and the new data set.
# Merge them and then exclude duplicates.
# Print some info about this

print(paste("Rows in current data set: ", nrow(df_data)))
print(paste("Rows in new data set: ", nrow(df_new)))

df_merge <- bind_rows(df_data, df_new)

print(paste("After merging: ", nrow(df_merge)))

df_merge <- distinct(df_merge)
print(paste("Total number of distinct rows in merged dataframe: ",  nrow(df_merge)))

```




```{r export-data}

df_merge %>% 
    select(dc_isodate, dc_year, dc_month, delay_mins, dest, dest_name, dc_reason, treat_mins, total_mins) %>% 
    write_csv(filename_out)


```

